---
title: "Assignment-07"
author: "Joshua Wade"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, error = FALSE)
library(tidyverse)
library(kableExtra)
library(HelpersMG)
library(choroplethr)
library(choroplethrMaps)
library(dplyr)
library(ggplot2)
```

```{r Step 3, message = FALSE, echo = FALSE}
Data1 <- read_csv("US_MAP_DATA.csv", skip = 2)
deathData <- Data1 %>% 
        rename(Deaths = "Total Death") %>%
        filter(fips < 57, abbr != "US") %>%
        summarise(region = str_to_lower(jurisdiction), value = Deaths)
write.csv(deathData, "Assignment-07", row.names = FALSE)
```

```{r Step 4, message = FALSE, echo = FALSE}
rateData <- Data1 %>% 
        filter(fips < 57, abbr != "US") %>%
        summarize(region = str_to_lower(jurisdiction), value = RatePer100000)
write.csv(rateData,"Assignment-07", row.names = FALSE)
```


```{r Step 5, message = FALSE, echo = FALSE}
state_choropleth(deathData, title = "Total Deaths due to COVID", legend = "Deaths")
```
<br/>
<br/>

### Interpretation
Covid deaths seems to follow a trend of higher populated areas are more affected, which makes sense considering it spreads most commonly though close proximity, this means that more rural areas are less likely to get it, or it could boil down to other factors like rural areas have lower rates due to lack of modernized medicine due to a broken health care system, but i digress

<br/>
<br/>
<br/>
<br/>
<br/>
<br/>
```{r step 6, message = FALSE, echo = FALSE}
state_choropleth(rateData, title = "Total Deaths Normalized", legend = "Deaths")
```
<br/>
<br/>

### Interpretation 

When it comes to Covid deaths vs Normal deaths Covid deaths in some areas are almost 4x as much as normalized deaths making it the most common death for a person to have. this is most likely due to the close proximity of these areas with inner cities and suburbs make these areas harder to control. it also could be due to lack of testing in certain places.
